Dialect: A Compact Line Notation Based on SMILES

### 2022-05-25 manuscript.tex
Drops support for encoding zero for hydrogens, charge, and isotope attributes. closes #104

\begin{table}
\begin{minipage}{\textwidth}
\renewcommand*\footnoterule{}
\caption{Data Model Primitives.}
\label{table:primitives}
\centering
\begin{tabular}{l l l}
    \hline
    Name & Description & Values\footnote{Ranges defined as \ttt{a..b} include \ttt{a} and exclude \ttt{b}.} \\
    \hline
    \ttt{Option<Type>} & Optional Value & \ttt{None}, \ttt{Type} \\
    \ttt{Index} & Unsigned integer & 0..2\textsuperscript{32} \\
    \ttt{Three} & Integer count & 1..4 \\
    \ttt{Ten} & Unsigned integer & 0..10 \\
    \ttt{SignedTen} & Integer count & -9..10 \\
    \ttt{Thousand} & Integer count & 0..1000 \\
    \ttt{Element} & IUPAC-\-approved element symbol & \ttt{Ac}, \ttt{Ag}, \ttt{Al}, {\ldots}\ttt{Zn}\\
    \ttt{HydrogenCount} & Hydrogen count & \ttt{Implicit}, \ttt{Ten} \\
    \ttt{Configuration} & Configurational descriptor & \ttt{Right}, \ttt{Left} \\
    \ttt{Direction} & Conformational descriptor & \ttt{Up}, \ttt{Down} \\
    \ttt{boolean} & Boolean & \ttt{true}, \ttt{false} \\
    \hline
\end{tabular}
\end{minipage}
\end{table}

@@

\begin{table}
\caption{Atomic State.}
\centering
\begin{tabular}{l L L l}
    \hline
    Attribute & Description & Type & Default \\
    \hline
    \ttt{index} & Unique identifier &\ttt{Index} & - \\
    \ttt{element} & Elemental symbol & \ttt{Option} \ttt{<Element>} & \ttt{None} \\
    \ttt{hydrogens} & Hydrogen count & \ttt{HydrogenCount} & 0 \\
    \ttt{configuration} & Configuration & \ttt{Configuration} & \ttt{None} \\
    \ttt{charge} & Formal charge & \ttt{SignedTen} & 0 \\
    \ttt{isotope} & Mass number & \ttt{Option<Thousand>} & \ttt{None} \\
    \ttt{selected} & Selection status & \ttt{boolean} & \ttt{false} \\
    \hline
\end{tabular}
\label{table:atomic-state}
\end{table}

@@

The attributes \ttt{element} and \ttt{isotope} characterize an atom's nucleus. The \ttt{element} attribute is an optional one- or two-letter character sequence selected from the set designated by the IUPAC~\cite{periodicTable}. When IUPAC authorizes additional symbols, they will become valid values for the \ttt{element} attribute. If the elemental identity of an atom is unknown, its \ttt{element} attribute must equal \ttt{None}. The \ttt{isotope} attribute is an optional integer value representing an atom's nuclear mass number, where mass number is the sum of proton and neutron count. Setting the \ttt{isotope} property to \ttt{None} means that the element's isotopic composition equals natural abundance. Otherwise, the lower bound on the value of \ttt{isotope} is zero. This lower limit allows for such physically nonsensical states such as atoms of zero implied mass (\ttt{isotope} equals zero) or negative implied neutron count (e.g., carbon-5). Implementations may reject such atomic states as invalid.

@@

The value of a bracket atom's \ttt{isotope} attribute is determined by the optional \ttt{<isotope>} non-terminal. It consists of between one and three digits encoding the integers 0-999. Leading zeros are disregarded when assigning the value of the \ttt{isotope} attribute. In other words, \enquote{007} and \enquote{7} are considered equivalent expressions of the value 7.

\begin{lstlisting}
<isotope> ::= <digit> <digit>? <digit>?
\end{lstlisting}

@@

\begin{lstlisting}
<virtual_hydrogen> ::= "H" <digit>?
\end{lstlisting}

The \ttt{<charge>} production rule sets the \ttt{charge} attribute of a bracket atom. This non-terminal begins with either the plus or minus terminals (\enquote{+} and \enquote{-} respectively) and ends with an optional \ttt{<digit>} non-terminal. A missing digit causes the atomic \ttt{charge} attribute to be set to one. In other words, the plus terminal (\enquote{+}) sets the \ttt{charge} attribute to one and the minus terminal (\enquote{-}) sets it to minus one. A charge of zero (e.g., \enquote{+0} or \enquote{-0}) sets the \ttt{charge} attribute to zero.

\begin{lstlisting}
<charge> ::= ( "+" | "-" ) <digit>?
\end{lstlisting}

@@

The \ttt{<cut>} non-terminal (\enquote{cut}) can takes two forms. A digit non-terminal can be used, enabling single-digit cut indexes. Double- or single-digit cut indexes, supporting the values zero through 99 inclusive, are available by prepending the percent character (\ttt{\%}). For the latter variant, any leading zero is disregarded, meaning that the sequences "07" and "7" are both interpreted as the index seven.

\begin{lstlisting}
<cut> ::= <digit> | "%" <digit> <digit>
\end{lstlisting}

>>>>

\begin{table}
\begin{minipage}{\textwidth}
\renewcommand*\footnoterule{}
\caption{Data Model Primitives.}
\label{table:primitives}
\centering
\begin{tabular}{l l l}
    \hline
    Name & Description & Values\footnote{Ranges defined as \ttt{a..b} include \ttt{a} and exclude \ttt{b}.} \\
    \hline
    \ttt{Option<T>} & Optional Value & \ttt{None}, \ttt{T} \\
    \ttt{Index} & Unsigned integer & 0..2\textsuperscript{32} \\
    \ttt{Three} & Integer count & 1..4 \\
    \ttt{Dec} & Integer count &0..10 \\
    \ttt{SignedDec} & Integer count & -9..10 \\
    \ttt{Thou} & Integer count & 1..1000 \\
    \ttt{Element} & IUPAC-\-approved element symbol & \ttt{Ac}, \ttt{Ag}, \ttt{Al}, {\ldots}\ttt{Zn}\\
    \ttt{Configuration} & Configurational descriptor & \ttt{Right}, \ttt{Left} \\
    \ttt{Direction} & Conformational descriptor & \ttt{Up}, \ttt{Down} \\
    \ttt{Implicit} & Implicit hydrogen count flag & - \\
    \ttt{boolean} & Boolean & \ttt{true}, \ttt{false} \\
    \hline
\end{tabular}
\end{minipage}
\end{table}

@@

\begin{table}
\caption{Atomic State.}
\centering
\begin{tabular}{l L L l}
    \hline
    Attribute & Description & Type & Default \\
    \hline
    \ttt{index} & Unique identifier &\ttt{Index} & - \\
    \ttt{element} & Elemental symbol & \ttt{Option} \ttt{<Element>} & \ttt{None} \\
    \ttt{hydrogens} & Hydrogen count & \ttt{Implicit}, \ttt{Dec} & 0 \\
    \ttt{configuration} & Configuration & \ttt{Option} \ttt{<Configuration>} & \ttt{None} \\
    \ttt{charge} & Formal charge & \ttt{SignedDec} & 0 \\
    \ttt{isotope} & Mass number & \ttt{Option<Thou>} & \ttt{None} \\
    \ttt{selected} & Selection status & \ttt{boolean} & \ttt{false} \\
    \hline
\end{tabular}
\label{table:atomic-state}
\end{table}

@@

The attributes \ttt{element} and \ttt{isotope} characterize an atom's nucleus. The \ttt{element} attribute is an optional one- or two-letter character sequence selected from the set designated by the IUPAC~\cite{periodicTable}. When IUPAC authorizes additional symbols, they will become valid values for the \ttt{element} attribute. If the elemental identity of an atom is unknown, its \ttt{element} attribute must equal \ttt{None}. The \ttt{isotope} attribute is an optional integer value representing an atom's nuclear mass number, where mass number is the sum of proton and neutron count. Setting the \ttt{isotope} property to \ttt{None} means that the element's isotopic composition equals natural abundance. Otherwise, the lower bound on the value of \ttt{isotope} is one. This lower limit allows for such physically nonsensical states such as atoms of negative implied neutron count (e.g., carbon-5). Implementations may reject such atomic states as invalid.

@@

The value of a bracket atom's \ttt{isotope} attribute is determined by the optional \ttt{<isotope>} non-terminal. It consists of between one and three digits encoding the integers 1-999. Leading zeros (e.g., \enquote{007} are disallowed.

\begin{lstlisting}
<isotope> ::= <nonzero> <digit>? <digit>?
\end{lstlisting}

@@

\begin{lstlisting}
<virtual_hydrogen> ::= "H" <nonzero>?
\end{lstlisting}

The \ttt{<charge>} production rule sets the \ttt{charge} attribute of a bracket atom. This non-terminal begins with either the plus or minus terminals (\enquote{+} and \enquote{-} respectively) and ends with an optional \ttt{<digit>} non-terminal. A missing digit causes the atomic \ttt{charge} attribute to be set to one. In other words, the plus terminal (\enquote{+}) sets the \ttt{charge} attribute to one and the minus terminal (\enquote{-}) sets it to minus one. Charge values of zero are obtained if \ttt{<charge>} production is present.

\begin{lstlisting}
<charge> ::= ( "+" | "-" ) <nonzero>?
\end{lstlisting}

@@

The \ttt{<cut>} non-terminal (\enquote{cut}) can takes two forms. A digit non-terminal can be used, enabling single-digit cut indexes in the range one through nine, inclusive. Double- or single-digit cut indexes, supporting the values ten through 99 inclusive, are available by prepending the percent character (\ttt{\%}). For the latter variant, leading zeros are disallowed, meaning that the sequence \enquote{\%07} must generate an error.

\begin{lstlisting}
<cut> ::= <nonzero> | "%" <nonzero> <digit>
\end{lstlisting}
